---
title: "Multivariate Assignment"
author: "Aharon V. Zúñiga"
date: "10/19/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

<!-- > round(pcaraw$sdev^2,2) -->
<!--  [1] 0.70 0.27 0.12 0.08 0.03 0.03 0.02 0.02 0.01 0.01 -->
<!-- > 0.83789818^2 -->
<!-- [1] 0.7020734 -->
<!-- > round(0.83789818^2,2) -->
<!-- [1] 0.7 -->
<!-- > sqrt(0.7020734) -->
<!-- [1] 0.8378982 -->



```{r setup, include=FALSE}
#Libaries 
#install.packages(dplyr)
#install.packages(Matrix.utils)
#install.packages("maptools")
library(maptools)
library("dplyr")
library("Matrix.utils")
library(ggplot2)
library("lavaan")
knitr::opts_chunk$set(echo = TRUE)
```

# 0 Descriptive analysis. :D

1 V_creative It is important to this person to think up new ideas and be creative; to do things one’s own way.
2 V_rich It is important to this person to be rich; to have a lot of money and expensive things.
3 V_secure Living in secure surroundings is important to this person; to avoid anything that might be dangerous.
4 V_spoil_oneself It is important to this person to have a good time;to “spoil” oneself. 
5 V_do_good It is important to this person to do something for the good of 
6 V_be_successful Being very successful is important to this person; to have people recognize one’s achievements.
7 V_exciting_life Adventure and taking risks are important to this person; to have an exciting life.
8 V_behave_properly It is important to this person to always behave properly; to avoid doing anything people would say is wrong.
9 V_protect_environment Looking after the environment is important to this person; to care for nature and save life resources.
10 V_tradition Tradition is important to this person; to follow the customs handed down by one’s religion or family


```{r step_1}
load("~/Documents/Multivariante/R_Scripts/wvs.Rdata")
summary(wvs[,1:10])
zscore <- function(x) {
  zscore <- (x-mean(x))/sd(x)
  } 
```

# 1. PCA + Biplot. 

### here explanation  (empirical analysis)

#### Decision, the first component will have the maximun variance of the raw data. 
```{r step_2}
zgspraw <- apply(wvs[,1:10],2,zscore) 
zgspraw <- cbind(zgspraw,country=wvs$country)
summary(zgspraw)
```

```{r step_3}

av <- mean(zgspraw[,1])
c<-aggregate.Matrix(zgspraw, groupings = zgspraw[,'country',drop=FALSE],fun='mean')
c <-as.matrix(c)

```

```{r step_4}

pcaraw <-prcomp(c[,1:10]) 
attributes(pcaraw)
#variance
round(pcaraw$sdev^2,2)
# proportion of explained variance per component ****
round(pcaraw$sdev^2/10,5)

# compute component loadings #pcaraw$rotation
A<-pcaraw$rotation%*%diag(pcaraw$sdev) #%*% multiplication matrix

#loading multiplied by variance of each variable
plot(A[,1:2],xlab="PC1",ylab="PC2")

pointLabel(A[,1],A[,2],names(wvs[,1:10]),
   cex=1) 

####
# compute unstandardized component scores
Zun<-zgspraw[,1:10]%*%pcaraw$rotation[,1:10]
# compute standardized component scores
Zs<-zgspraw[,1:10]%*%pcaraw$rotation[,1:10]%*%diag(1/pcaraw$sdev[1:10])
# plot component scores for first 2 unstandardized components
plot(Zun[,1:2],xlim=c(-3,8),ylim=c(-3,3),xlab="PC1",ylab="PC2")
#pointLabel(Zun[,1],Zun[,2],wvs[,10],cex=0.9) 

```






# 2. Exploratory Factor Analysis :D

column	variable	item									
16. J_claiming_benefits	Claiming government benefits to which you are not entitled  
17.	J_avoiding_fare	Avoiding	a	fare	on	public	transport  
18.	J_stealing_property	Stealing	property  
19.	J_cheating_taxes	Cheating	on	taxes	if	you	have	a	chance  
20.	J_accept_bribe	Someone	accepting	a	bribe	in	the	course	of	their	duties  
21.	J_homosexuality	Homosexuality  
22.	J_prostitution	Prostitution  
23.	J_abortion	Abortion  
24.	J_divorce	Divorce  
25.	J_sex_before_marriage	Sex	before	marriage  
26.	J_suicide	Suicide  
27.	J_beat_wif	For	a	man	to	beat	his	wife  
28.	J_parents_beating_children	Parents	beating	children  
29.	J_violence	Violence	against	other	people  



```{r 2step_1}
#print matrix of observed correlations

zjustif <- apply(wvs[,16:29],2,zscore) 
cormat <- round(cor(zjustif),2)
covmat <- round(cov(zjustif),2)

#round(cov(wvs[,16:29]),2) when data is not stadarized, we see the differences.


## estimate EFA with xxx common factors
## The factanal() function runs EFA on the correlation matrix
fa_justif <-factanal(zjustif, factors=7, rotation="none")
fa_justif


#EFA with varimax rotation and 3 factors - varimax orthogonal - prohibited cor.
favm_justif<-factanal(zjustif,factors=3, rotation="varimax")
print(favm_justif, digits = 3, cutoff = 0.4, sort = TRUE)

#EFA with promax rotation and 4 factors - promax oblique - allows correlation between factors
favm_justif2 <- factanal(zjustif,factors=4, rotation="promax")
print(favm_justif2, digits = 3, cutoff = 0.4, sort = TRUE)

pjustif<-prcomp(zjustif)
screeplot(pjustif,type="lines")


## score -- I choose towards to find a explanation in 3 areas. Financial issues
## interpersonal issues and Violence. 
favm_justif<-factanal(zjustif,factors=3, rotation="varimax",scores="regression")
print(favm_justif, digits = 3, cutoff = 0.5, sort = TRUE)
scores <- favm_justif$scores
colnames(scores)<-c("Interpersonal","Financial", "Violence")
zjustif_score <- data.frame(zjustif,scores,country=wvs$country)

ggplot(data = zjustif_score,aes(Interpersonal, fill=country)) +
  geom_density()  + 
 facet_wrap(~country) + theme(legend.position = "none") +
  ggtitle("Factor 1 - Interpersonal Justification")
  
ggplot(data = zjustif_score,aes(Financial, fill=country)) +
  geom_density()  + 
 facet_wrap(~country) + theme(legend.position = "none") +
  ggtitle("Factor 2 - Financial Justification")

ggplot(data = zjustif_score,aes(Violence, fill=country)) +
  geom_density()  + 
 facet_wrap(~country) + theme(legend.position = "none") +
  ggtitle("Factor 3 - Violence Justification")

##### standarise.

```


# 3. Confirmatory Factor Analysis :)

```{r 3step_1}
#pas <- wvs[wvs$country %in% c("Netherlands","Malaysia"),]
#tail(pas)
zjustif_cfa <- apply(wvs[wvs$country %in% c("Netherlands","Malaysia"),16:29],2,zscore)
covmat_cfa <- cov(zjustif_cfa) #no round to get more detailed data
~
cfa3 <- 'Interpersonal =~ 1*J_homosexuality+J_prostitution+J_abortion+J_divorce+J_sex_before_marriage+J_suicide
   Financial =~ 1*J_claiming_benefits+J_avoiding_fare+J_stealing_property+J_cheating_taxes+J_accept_bribe
   Violence =~ 1*J_beat_wife+ J_parents_beating_children+J_violence
   Financial ~~ Violence
   Violence =~ J_suicide
   Violence =~ J_prostitution'

fitcfa3 <- cfa(cfa3, sample.cov = covmat_cfa, sample.nobs = 2473)

summary(fitcfa3, fit.measures=TRUE)
standardizedSolution(fitcfa3)
modificationIndices(fitcfa3, minimum.value = 10, sort = TRUE)


semPaths(fitcfa3, title = FALSE, curvePivot = TRUE)

semPaths(survey.scalfit3,"model","std","lisrel", edge.label.cex = 1.2, intercepts = FALSE, layout = "tree2",
                     panelGroups = FALSE, ask  = FALSE, groups = "latent", pastel = TRUE, exoCov = TRUE, rotation = 1)gm




